---
title: "Data Appendix to \"The impact of weather events on schooling outcomes\""
author: "Katya Garcia-Israel"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F, 
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

# add haven package to make stata 16 version readable 
library(foreign)
library(haven)

#The following custom function simplifies the process of writing dfSummaries to html files
export_summary_table <- function(dfSummary_output){
  data_info <- attr(dfSummary_output, "data_info")
  ds_name <- data_info$Data.frame
  print(dfSummary_output,
      file = str_c("output/", ds_name, "_summary.html"),
      method = "browser",
      report.title = ds_name)
}
```

```{r set dfSummary css style, echo = F, include = F}
st_css()
```
# Appendix description
*Your Data Appendix should begin with a brief statement explaining its purpose like the following one.*

The datasets used directly by the final analysis are saved in `processed-data/` at the end of this file.

# Raw data
This section documents the datasets used in this analysis.

## Household Survey data
**Citation:** .  
Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales

**Date Downloaded:** Downloaded September 2019.
**Filename(s):** raw_data/filename.csv 
**Unit of observation:** Unit of observation at the individual level, but each individual is identified mainly by the household id. 
**Dates covered:** 2005 

### To obtain a copy

This household survey can be found on the Bolivian National Statistical Institute website, under the menu for 'Banco de Datos' and 'Base de datos de encuestas sociales.' Then the household surveys may be found for the years 2005-2018, excepting 2010.

The original format of this data was from the sav file, but I imported it into Stata before importing the Stata format into R.

### Variable descriptions

- **hh_id:** The household ID, which is the same for each member of the household. It is made up of letters and numbers. 
- **nro1:** The number of the person in each household.
- **assist:** Binary variable, whether or not the individual assists school 
- **urb_rur:** Binary variable, whether the household is in an urban or rural area
- **reason:** the reason why the individual is not assisting school if they are not 
- **members:** number of household members excluding servants 
- **sex:** the sex of the individual, Male = 1, Female =2 
- **age:** How old the individual is, in years 
- **relation:** The relationship of the individual to the head of household 
- **literacy:** Binary variable, whether a person can read and write 
- **register:** Binary variable, whether an individual registered for school or not 
- **grade:** For what grade is the person registered 
- **repeated:** whether the person has repeated the grade they are currently registered in 
- **dept:** the department in which the person lives 


### Data import code and summary

```{r import household data}
data_2005 <- read_dta(file.path("Raw data","seminar_paper_data","2005_hhsurvey.dta")) %>% 
  rename(hh_id = folio,
        assist = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy = s3_01,
        register = s3_04,
        grade = s3_05b,
        repeated = s3_07) %>% 
  mutate(dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"))

hh_2005 <- data_2005 %>% 
  select(hh_id, assist, reason, members, sex, age, relation, literacy, register, grade, repeated, dept, urb_rur)
```



```{r exporting data table}
export_summary_table(dfSummary(hh_2005))
```

## Flood data
**Citation:** GeoSinager. (2016). Mapa Recurrencias de Inundaciones Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/226

Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales

**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
**Date Downloaded:** Identify when you downloaded the dataset.  
**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
**Unit of observation:** What distinguishes different rows in your dataset?  
**Dates covered:** What time frame does the data cover?  

### To obtain a copy

Describe in a step-by-step fashion how an interested user could obtain the data.

### Importable version (if necessary)

**Filename(s):** importable-data/filename-importable.csv

In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one. 

### Variable descriptions

Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.

- **dept:** department name 
- **A_2005:** how many floods occurred in 2005 in that location 
- **A_2006:** how many floods occurred in 2006 in that location 
- **A_2007:** how many floods occurred in 2007 in that location 
- **A_2008:** how many floods occurred in 2008 in that location 
- **A_2009:** how many floods occurred in 2009 in that location 
- **A_2010:** how many floods occurred in 2010 in that location 
- **A_2011:** how many floods occurred in 2011 in that location 
- **A_2012:** how many floods occurred in 2012 in that location 

### Data import code and summary
```{r import flood data}
flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
  rename(dept = NOM_DEP)

flood_summary <- flood %>% 
  group_by(dept) %>% 
  summarize_at(vars(starts_with("A_")), sum) %>% 
  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "floods")

export_summary_table(dfSummary(flood_summary))
```
## Drought data
**Citation:** GeoSinager. (2016). Mapa Recurrencias de Sequias Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/225

**Date Downloaded:** October 2019.  
**Filename(s):** Raw data/droughts.csv 
**Unit of observation:** The unit of observation is the municipality 
**Dates covered:** 2002-2012  

### To obtain a copy

This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225. It is shown on a map with different layers. 

### Importable version (if necessary)

**Filename(s):** importable-data/filename-importable.csv

In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one. 

### Variable descriptions

Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.

- **dept:** department name 
- **A_2005:** how many droughts occurred in 2005 in that location 
- **A_2006:** how many droughts occurred in 2006 in that location 
- **A_2007:** how many droughts occurred in 2007 in that location 
- **A_2008:** how many droughts occurred in 2008 in that location 
- **A_2009:** how many droughts occurred in 2009 in that location 
- **A_2010:** how many droughts occurred in 2010 in that location 
- **A_2011:** how many droughts occurred in 2011 in that location 
- **A_2012:** how many droughts occurred in 2012 in that location 


```{r importing droughts data, echo = F, message = F}
#drought data set 
library(readxl)
drought <- read_excel(file.path("Raw data", "recurrenciasequia_2002_2012bol (1).xls")) %>% 
#rename variables
    rename(dept = NOM_DEP)

drought_summary <- drought %>%
  group_by(dept) %>% 
  summarize_at(vars(starts_with("A_")), sum) %>%
  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "droughts")

export_summary_table(dfSummary(drought_summary))

```

# Data Processing and Combination
To create a complete dataset, I merged the household data with the floods data and the droughts data. Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis.
*This section should include a discussion of the processing and merging steps needed to create your basic data. The code to implement these steps should be included in chunks in this section. Once the final merged data has been created, you should use the dfSummary function again to summarize the data you will be using. You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.*

```{r merging datasets}
#I would recommend combining all the years of household data and then merging with the weather data

total_2005 <- hh_2005 %>% 
  left_join(flood_summary) %>%
  left_join(drought_summary)
```

# Analysis Variables

This section should include a description of all the variables that are used in your final analysis. At the end of the section, you should save all of these variables in the processed_data folder of your repository.

# Discussion of Data

In the household data, the sample of children in 2005 was 

*This section should include a discussion of any data patterns you notice based on the summaries created in the code above.*
