---
title: "The Effect of Weather Shocks on Children's Education in Bolivia"
author: "Katya Garcia-Israel"
abstract: "Weather shocks have the potential to dramatically affect households' livelihoods and income levels, especially in developing countries. A growing subset of developmental and environmental literature focuses on investigating the relationship between environment and weather factors and household changes in behavior. One frequently cited response to negative weather shocks that decrease income is an increase in child labor to supplement the household income and a decrease in school attendance. This mechanism is especially relevant in agricultural areas, but can also be seen in more urban areas if the weather events cause a loss of property, injury, or displace people. This study explores the relationship between flood and drought occurrences and school attendance in Bolivia. I find that ..."
header-includes:
   - \usepackage{booktabs}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage[labelfont=bf]{caption}
output: 
  pdf_document:
    number_sections: true
    keep_tex: true

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F, 
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

# add haven package to make stata 16 version readable 
library(foreign)
library(haven)

#The following custom function simplifies the process of writing dfSummaries to html files
export_summary_table <- function(dfSummary_output){
  data_info <- attr(dfSummary_output, "data_info")
  ds_name <- data_info$Data.frame
  print(dfSummary_output,
      file = str_c("output/", ds_name, "_summary.html"),
      method = "browser",
      report.title = ds_name)
}
```
# Introduction


# Literature Review

Several papers have researched the effects of weather events on general household well-being, examining household responses to shocks on income. Baez et al (2017) find that rural households in Guatemala responded to income shocks after a tropical storm by increasing child labor. Such responses are negatively related to school attendance. Jensen (2000) examines differences in school enrollment rates after rainfall shocks and compares them to unaffected areas, finding that investments in children’s education do decrease in areas that have experienced either heavy or low rainfall.

The impacts of weather shocks could potentially be seen some time after their occurrence. Marchetta et al (2019) address this by including lagged effects in their model, using a weather variable lagged by time to show the potential for a delayed effect on child outcomes. The reasoning behind this, in their case, was that households would observe the impact of the rainy season prior to enrolling in school before making the decision to send their children to school or not. Marchetta et al (2019) seek to answer the question of how the duration of the negative effects of weather shocks on education by testing how test scores are impacted after the shock. Instead of measuring only school attendance, including test scores in the analysis allows for the possibility of seeing patterns in lowered test scores after the weather event. This again suggests the potential of lagged impacts.Shah and Steinberg (2017) found differing effects on enrollment from distinct weather events; droughts led to an increase in school attendance and floods decreased it. The authors gave a few different explanations for why they might be finding these results. Rainfall could be linked to increased malaria and diarrhea which would keep children from attending school or the school meals being served more during drought seasons could serve as an incentive for attending school.

Negative weather shocks can impact different groups in distinct ways. For instance, Marchetta et al find that girls are more negatively affected by extreme weather events. In research on India, Zimmermann (2020) also shows that rainfall shocks affect girls’ schooling more than boys’, and younger children are more insulated from the effects of rainfall shock than older children who are more productive workers. This preference for educating boys does not hold in all contexts; Jensen (2000) finds no difference between boys’ and girls’ school enrollment as a result of rainfall shocks in Cote d’Ivoire.

As opposed to examining the mechanisms of decreased incomes through damaged harvests, Groppo and Kraehnert (2017) explore the effects of livestock mortality on household incomes and education in Mongolia. Their study suggests that weather shock of harsh winters does not affect long-term education choices, as household income decreases but the children do not substitute schooling for labor. The mechanisms by which weather events affect child outcomes are not always tied directly to agriculture; Marchetta et al (2017) study cyclones as well as rainfall variability, which suggests that loss of (capital) and damage to property, as well as the larger economy, could play a role in the household response that is not necessarily tied to decreasing income from harm to agriculture. Their work also emphasizes lack of access to insurance and savings as a main determinant of the negative child outcomes. Teacher attendance during drought rather than rainfall seasons and impacts of migration were other possible mechanisms discussed by Shah and Steinberg (2017).

The methods used in this literature vary. Baez et al (2017) used a regression-based analysis of household consumption and different levels of rainfall varying by location, as well as setting up a probability model of which areas would be most affected by rainfall as a robustness check. Marchetta et al (2019) used a bivariate probit model as their main empirical strategy and a linear probability model as a robustness check. Groppo and Kraehnert (2017) use a difference-in-differences approach to distinguish between areas affected by weather disasters and those that were not. Jensen (2000) examines the differences between enrollment rates in areas that experienced a heavy rainfall shock and areas that did not, using deviation from the long-term mean as an indication of a shock. The identifying assumption is made that there would have been no differential treatment of children’s schooling between geographic areas in Cote d’Ivoire without the negative rainfall shock, an assumption that is supported by examining the percentages of children enrolled in a prior year when neither area experienced a weather shock. Researchers are faced with the problem of ensuring to the best of their ability that the observed changes in household outcomes are not due to unobserved changes unrelated to weather patterns.



# Data and Methods

## Conceptual Framework
I attempt to answer the question of how weather events impact school attendance and registration using an empirical approach. The theoretical arguments for the analysis are based on the ideas discussed in the literature review, that extreme weather events can impact households’ financial situations, especially in areas where agriculture is important for the local economy and people’s livelihoods. The connection to schooling comes from the idea that financial burdens can cause children to substitute the time they would have spent in school into participating in economic activities to help increase household income. In rural areas, this effect is expected to be the strongest, as there is larger participation in agriculture and many households may use all their members at one point or another to help with the agricultural activities. However, both the urban and rural samples are relevant to the research question because agricultural activity might be taking place in some areas classified as urban. Additionally, other mechanisms linking floods and droughts to school attendance may be at play, such as decreases in real income from property damage, livestock mortality, or higher food prices.

## Data
To answer this question, I used a combination of household survey data and flood and drought occurrences data. The household survey data was taken from the Bolivian National Statistical Institute as part of their social surveys database. The household surveys used were from the years 2005-2007 and 2011-2012. The intermediary years (2008, 2009, 2010) were not included because the relevant education questions were not part of the 2008 and 2009 surveys, and there was no household survey taken in 2010, the census year. The flood and drought datasets show the number of floods and droughts that occurred per year in each of the nine departments of Bolivia.  The final compiled dataset is a repeated cross-section as different households were randomly surveyed each year. The main unit of observation is the child’s ID, but the household ID is also used as a unit in order to examine household characteristics such as relations and income per person. For all the combined years, there are 27,921 observations. Table 1 lists the variables used to answer the research question, along with brief descriptions.


```{r import household data 2005, echo = F, message = F}
data_2005 <- read_dta("Raw data/seminar_paper_data/2005_hhsurvey.dta") %>% 
  rename(hh_id = folio,
        assist1 = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy1 = s3_01,
        register1 = s3_04,
        school_level = s3_05a,
        grade = s3_05b,
        repeated = s3_07,
        income = y_perc1A,
        educ = aoesc,
        job = s4_16cod,
        tasks = s4_16b,
        pov_line = z,
        extr_pov = z_ext) 

hh_2005 <- data_2005 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2005,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
         reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
              
```


```{r import household data 2006, echo = F, message = F}
data_2006 <- read_dta("Raw data/seminar_paper_data/2006_hhsurvey.dta")  %>% 
  rename(hh_id = Folio,
        assist1 = s4_08,  
        reason = S4_09, 
        members = mhogar,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy1 = s4_01,
        register1 = s4_04,
        school_level = s4_05a,
        grade = s4_05b,
        repeated = s4_07,
        income = yhogpcf,
        urb_rur = Urb_Rur,
        depto = ID01,
        educ = añoe,
        job = s5_17cod,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext) 

hh_2006 <- data_2006 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2006,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
```

```{r import household data 2007, echo = F, message = F}
data_2007 <- read_dta("Raw data/seminar_paper_data/2007_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_10,  
        reason = s4_11, 
        members = mhogar,
        sex = s1_03,
        age = s1_04,
        relation = s1_06,
        literacy1 = s4_01,
        register1 = s4_06,
        school_level = s4_07a,
        grade = s4_07b,
        repeated = s4_09,
        income = yhogpcf,
        voucher = s4_05,
        educ = a_oe,
        job = s5_16co,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext) 

hh_2007 <- data_2007 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2007,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
```

```{r import household data 2011, echo = F, message = F}
data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_05,
        grade = s4_06b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_06a,
        urb_rur = area,
        breakfast = s4_07,
        educ = e,
        job = cod_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

hh_2011 <- data_2011 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2011,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                               literacy1 == 2 ~ 0), 
          female = case_when(sex == 2 ~ 1,
                             sex == 1 ~ 0),
           rural = case_when(urb_rur == 1 ~ 0,
                             urb_rur == 2 ~ 1), 
           attend = case_when(assist1 == 1 ~ 1,
                              assist1 == 2 ~ 0),
           register = case_when(register1 == 1 ~ 1,
                                register1 == 2 ~ 0),
         primary = case_when(school_level == 16 ~ 1,
                             school_level != 16 ~ 0),
         secondary = case_when(school_level == 17 ~ 1,
                               school_level != 17 ~ 0))
```

```{r household data 2012, echo = F, message = F}
data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_04,
        grade = s4_05b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_05a,
        urb_rur = area,
        depto = departamento,
        breakfast = s4_07,
        voucher = s4_08,
        voucher_spent = s4_09,
        educ = e,
        job = cods5_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

hh_2012 <- data_2012 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2012,
         dept = case_when(depto== "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
         literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 19 ~ 1,
                             school_level != 19 ~ 0),
         secondary = case_when(school_level == 20 ~ 1,
                               school_level != 20 ~ 0))

```

```{r combine household datasets, echo = F, message = F}
all_years <- hh_2005 %>% 
  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012) 
 

```


```{r import flood data, echo = F, message = F}
flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
  rename(dept = NOM_DEP)

flood_summary <- flood %>% 
  group_by(dept) %>% 
  summarize_at(vars(starts_with("A_")), sum) %>% 
  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "floods") %>%
  mutate(year = as.numeric(year))

#lagged flood data created in excel
flood_lag <- read_csv(file.path("Raw data", "lagflood.csv"))
#library(DataCombine)
#flood_lags <- slide(flood_summary, Var = "floods", slideBy = -1)

flood_total <- flood_summary %>% 
  left_join(flood_lag)
  
```

```{r importing droughts data, echo = F, message = F}
#drought data set 
library(readxl)
drought <- read_excel(file.path("Raw data", "recurrenciasequia_2002_2012bol (1).xls")) %>% 
#rename variables
    rename(dept = NOM_DEP)

drought_summary <- drought %>%
  group_by(dept) %>% 
  summarize_at(vars(starts_with("A_")), sum) %>%
  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "droughts") %>%
  mutate(year = as.numeric(year))

#lagged drought data created in excel
#library(xlsx)
#write.xlsx(drought_summary, "C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\lagdrought.xlsx")

drought_lag <- read_csv(file.path("Raw data", "lagdrought.csv"))

drought_total <- drought_summary %>% 
  left_join(drought_lag)
  

export_summary_table(dfSummary(drought_total))

```

``` {r merging datasets, echo = F, message = F}
total_data<- all_years %>% 
  left_join(flood_total) %>%
  left_join(drought_total) %>%
  mutate(child = case_when(relation == 3 ~ 1,
                           relation != 3 ~ 0),
         grandchild = case_when(relation == 8 ~ 1,
                                relation != 8 ~ 0),
         pov_diff = income - pov_line,
         pov = case_when(pov_diff > 0 ~ 0,
                         pov_diff < 0 ~ 1),
         extr_diff = income - extr_pov,
         extr = case_when (extr_diff > 0 ~ 0,
                           extr_diff < 0 ~ 1),
                  work_reason = case_when(reason == "work" ~ 1,
                                 reason != "work" ~ 0),
         money_reason = case_when(reason == "money" ~ 1, 
                                  reason != "money" ~ 0))


#total_data %>%
  #group_by(work_reason, money_reason) %>%
  #summarise(n = n())

export_summary_table(dfSummary(total_data))


save("total_data", file = "processed_data/analysis_data.RData")

```


```{r creating head of household subset, echo = F, message = F}

hh_head <- subset(total_data, relation == 1, select = c("hh_id", "female", "age", "literacy", "dept", "rural", "income", "year", "educ", "job", "tasks", "pov_diff", "extr_diff", "pov", "extr"))

```


```{r creating children subset, echo = F, message = F}

children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "attend", "reason", "register", "primary", "secondary", "floods", "droughts", "lag_floods", "lag_droughts", "year", "money_reason", "work_reason"))

household_counts <- children %>%
group_by(hh_id) %>%
summarize(num_children = n())


children_id <- mutate(children, id = rownames(children))
  
total_children <- merge(children_id,household_counts, by="hh_id")

export_summary_table(dfSummary(total_children))


total_children %>%
      group_by(child, grandchild) %>%
      summarise(n = n())

total_children %>%
   group_by(work_reason, money_reason) %>%
   summarise(n = n())

```

```{r combining household head characteristics with children dataset, echo = F, message = F}
hh_head_2 <- hh_head %>%
  select(female, age, literacy, educ, job, tasks, income, pov, extr, hh_id) %>%
  rename (female_hhhead = female,
          hhhead_age = age,
          hhhead_literacy = literacy,
          hhhead_educ = educ,
          hhhead_job = job,
          hhhead_tasks = tasks)

total_heads_subset <- hh_head_2 %>% distinct()

#total <- total_children %>%
  #left_join(hh_head_2)
#left join command creating the duplicates problem it seems 
total <- merge(total_children,total_heads_subset, by="hh_id") 

total %>%
  group_by(reason) %>%
  summarise(n = n())

total_2 <- subset(total, !(reason == "vacation" & attend == 0)) #taking out the observations with vacation kids

total_subset <- total_2 %>% distinct()




```


```{r regression variable specifications, echo = F, message = F}

reg_data <- total_subset %>%
  mutate(year_factor = as.factor(year),
         dept_factor = as.factor(dept),
         lny = log(income + 1),
         valleys = case_when(dept %in% c("Chuquisaca", "Cochahamba", "Tarija") ~ 1, TRUE ~ 0),
         mountains = case_when(dept %in% c("La Paz", "Oruro", "Potosi") ~ 1, TRUE ~ 0),
         plains = case_when(dept %in% c("Beni", "Pando", "Santa Cruz") ~ 1, TRUE ~ 0),
         flood_secondary = floods * secondary,
         drought_secondary = droughts * secondary,
         flood_pov = floods * extr,
         drought_pov = droughts * extr,
         flood_lny = floods * lny,
         drought_lny = droughts * lny,
         lag_flood_lny = lag_floods * lny,
         lag_drought_lny = lag_droughts * lny,
         lag_f_secondary = lag_floods * secondary,
         lag_d_secondary = lag_droughts * secondary)

#adding the region variables for the region fixed-effects option
#adding interaction terms
#adding lagged flood and drought variables 

#creating rural sample
rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary", "lag_floods", "lag_droughts"))

reg_data %>%
  group_by(dept_factor) %>%
  summarise(n = n())


reg_data %>%
   group_by(attend, register) %>%
   summarise(n = n())

```
##Variable Descriptions

\begin{table}
\centering
\caption{Variable definitions}
\label{variableDefinitions}
\begin{tabular}{ll}
\hline
\hline
Variable name      & Description                                               \\
\hline
Attend          & Binary variable, 1 = attends school (of registered sample)   \\
Register        & Binary variable, 1 = registered for school                   \\
Work reason     & Binary variable, 1 = missing school for work                 \\
Lny             & Log of income per capita (in Bolivianos/month)               \\
Extr            & Binary variable, 1 = below extreme poverty line              \\
Female          & Binary variable, 1 = female                                  \\
Age             & Age in years                                                 \\
Child           & Binary variable, 1 = is child of head of household           \\
Grandchild      & Binary variable, 1 = is grandchild of head of household      \\
Num children    & Number of children in the household                          \\
Secondary       & Binary variable, 1 = in secondary school group               \\
hhhead educ     & Years of education of head of household                      \\
female hhhead   & Binary variable, 1 = female head of household                \\
Floods          & Number of flood occurrences per dept per year                \\
Droughts        & Number of drought occurrences per dept in current year       \\
Lag floods      & Number of flood occurrences per dept in the previous year    \\
Lag droughts    & Number of drought occurrences per dept in the previous year  \\
Dept            & Department (nine options)                                    \\
Valleys         & Binary variable, 1 = department is in valleys region         \\
Mountains       & Binary variable, 1 = department is in mountains region       \\
Plains          & Binary variable, 1 = department is in plains region          \\

\hline
\end{tabular}
\end{table}

```{r create summary statistic table, echo = F, results = 'asis'}
#creating a table with the main explanatory variables

table_total <- reg_data %>%
  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)
 

stargazer::stargazer(as.data.frame(table_total), 
                     summary = T,
                     label = "summaryStats",
                     title = "Summary statistics (full sample)",
                     header = F,
                     type = "latex",
                     digits = 3)

#creating a table dividing the characteristics by urban/rural area
table_rural <- rural %>%
  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains) 
  
stargazer::stargazer(as.data.frame(table_rural),
                     label = "summaryStatsR",
                     title = "Summary statistics (rural sample)",
                     summary = T,
                     header = F,
                     type = "latex",
                     digits = 3)

```

```{r creating line graphs for variation within flood variables, echo = F, message = F}

Beni <- subset(flood_summary, dept == "Beni" & year > 2003, select = c("year", "floods")) 
Chuq <- subset(flood_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "floods"))
Pando <- subset(flood_summary, dept == "Pando" & year > 2003, select = c("year", "floods"))
SantaCruz <- subset(flood_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "floods"))
Oruro <- subset(flood_summary, dept == "Oruro" & year > 2003, select = c("year", "floods"))
Potosi <- subset(flood_summary, dept == "Potosi" & year > 2003, select = c("year", "floods"))
LaPaz <- subset(flood_summary, dept == "La Paz" & year > 2003, select = c("year", "floods"))
Tarija <- subset(flood_summary, dept == "Tarija" & year > 2003, select = c("year", "floods"))
Cocha <- subset(flood_summary, dept == "Cochabamba" & year > 2003, select = c("year", "floods"))
# Create a first line
plot(LaPaz, type = "b", frame = FALSE, pch = 19, main = "Figure 2. Floods Over Time", cex.main = 1, col = "red", xlab = "year", ylab = "flood counts")
# Add lines
lines(Potosi, pch = 18, col = "orange", type = "b", lty = 1)
lines(Oruro, pch = 18, col = "orangered4", type = "b", lty = 1)
lines(Beni, pch = 18, col = "green", type = "b", lty = 1)
lines(Pando, pch = 18, col = "yellow", type = "b", lty = 1)
lines(SantaCruz, pch = 18, col = "green4", type = "b", lty = 1)
lines(Cocha, pch = 18, col = "midnightblue", type = "b", lty = 1)
lines(Tarija, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
lines(Chuq, pch = 18, col = "blue", type = "b", lty = 1)
# Add a legend to the plot
legend("topleft", legend=c("La Paz", "Potosi", "Oruro", "Beni", "Pando", "Santa Cruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)

```

```{r creating line graphs for variation within drought variables, echo = F, message = F}

Benid <- subset(drought_summary, dept == "Beni" & year > 2003, select = c("year", "droughts")) 
Chuqd <- subset(drought_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "droughts"))
Pandod <- subset(drought_summary, dept == "Pando" & year > 2003, select = c("year", "droughts"))
SantaCruzd <- subset(drought_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "droughts"))
Orurod <- subset(drought_summary, dept == "Oruro" & year > 2003, select = c("year", "droughts"))
Potosid <- subset(drought_summary, dept == "Potosi" & year > 2003, select = c("year", "droughts"))
LaPazd <- subset(drought_summary, dept == "La Paz" & year > 2003, select = c("year", "droughts"))
Tarijad <- subset(drought_summary, dept == "Tarija" & year > 2003, select = c("year", "droughts"))
Cochad <- subset(drought_summary, dept == "Cochabamba" & year > 2003, select = c("year", "droughts"))
# Create a first line
plot(LaPazd, type = "b", frame = FALSE, pch = 19, main = "Figure 3. Droughts Over Time", cex.main = 1, col = "red", xlab = "year", ylab = "drought counts")
# Add lines
lines(Potosid, pch = 18, col = "orange", type = "b", lty = 1)
lines(Orurod, pch = 18, col = "orangered4", type = "b", lty = 1)
lines(Benid, pch = 18, col = "green", type = "b", lty = 1)
lines(Pandod, pch = 18, col = "yellow", type = "b", lty = 1)
lines(SantaCruzd, pch = 18, col = "green4", type = "b", lty = 1)
lines(Cochad, pch = 18, col = "midnightblue", type = "b", lty = 1)
lines(Tarijad, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
lines(Chuqd, pch = 18, col = "blue", type = "b", lty = 1)
# Add a legend to the plot
legend("topleft", legend=c("La Paz", "Potosi", "Oruro", "Beni", "Pando", "Santa Cruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)

```

## Model Specification

The main econometric model used in the analysis is a probit regression model, as my three dependent variables are all binary variables with values of 0 or 1. Floods and droughts are the two main explanatory variables, along with the interaction terms floods_lny and droughts_lny, which capture the differing effect that weather shocks have on schooling outcomes for children of different income levels. These interaction terms were formed by multiplying the flood and drought variables with the lny, logged per capita monthly income, variable. Several control variables are inserted into the regression to control for income, age and sex of the child, sex and education level of the household head, number of other children in the household and whether the child is the child or grandchild of the household head. The subscript 'd' refers to the department and 't' to the year, as the probit regression utilizes both department-level and year fixed-effects. 

$1. attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$

$2. register_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$

$3. work\_reason_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$


In the first two regression equations, measuring school attendance and school registration, the omitted condition is not attending school and not registering for school, respectively. Thus, increases in the dependent variable indicate higher probabilities of going to school. Variables such as floods and droughts are expected to have negative coefficients, as I hypothesize that experiencing weather shocks will decrease the likelihood going to school. The age variable and the secondary school variable are both expected to have a negative coefficient, as older children would be more likely to be able to work and have a higher opportunity cost of going to school. A higher number of children in the household is expected to cause a decrease in the probability of attending school. Higher levels of income, being the child or grandchild of the household head, high levels of education of the household head, and having a female household head are all expected to be connected to higher probabilities of school attendance and registration. 

The third regression uses the sample of children that do not attend school, and the dependent variable is work_reason, which shows the likelihood of work being the reason why the child is not attending school. For the omitted condition of 0, those children listed reasons other than work as their motivation for not attending school. In this regression, I would expect the coefficients on the flood and drought variables to be positive, and the control variables to flip signs, as the factors that would cause a household to send a child to school have the opposite effect on the child’s likelihood of working. The sample for the third regression is much smaller than the other two regressions, 3,939 versus the 24,316 in the attend sample and the 27,921 in the registering sample.

These three regression equations are repeated in the analysis using different samples, the total number of households and the sample that live in rural areas. In the lagged variable analysis, the same model is used, but floods and droughts are replaced by lag_floods and lag_droughts, which represent the number of floods or droughts in each department during the previous year. Including the lagged variables is a way of testing whether households’ responses to weather shocks are long-term or short-term.


# Results



```{r probit regression total sample, echo = F, message = F}

probitMod_assist <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))  

probitMod_register <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

probitMod_work <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

```

```{r generate probit regression table, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
               title = "Full sample probit regression results", 
               label = "fullResults1",
               header = F,
               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

```

```{r probit regression rural sample, echo = F, message = F}

probitMod_assist_r <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))  

probitMod_register_r <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))

probitMod_work_r <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))

```

```{r generate probit regression table rural sample, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_r, probitMod_register_r, probitMod_work_r), 
               title = "Rural sample probit regression results", 
               label = "fullResults3",
               header = F,
               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

```

```{r probit regression urban sample, echo = F, message = F}

urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary", "lag_floods", "lag_droughts"))

probitMod_assist_u <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))  

probitMod_register_u <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))

probitMod_work_u <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))

```


The following tables show the results with the lagged variables for floods and droughts.
```{r probit regression total sample lagged variables, echo = F, message = F}

probitMod_assist_lt <- glm(attend ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))  

probitMod_register_lt <- glm(register ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

probitMod_work_lt <- glm(work_reason ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

```

```{r generate probit regression table lagged variables, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_lt, probitMod_register_lt, probitMod_work_lt), 
               title = "Full sample probit regression results (lagged)", 
               label = "fullResults5",
               header = F,
               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

```

```{r probit regression rural sample lagged, echo = F, message = F}

probitMod_assist_lr <- glm(attend ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))  

probitMod_register_lr <- glm(register ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))

probitMod_work_lr <- glm(work_reason ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))

```

```{r generate probit regression table rural sample lagged, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist_lr, probitMod_register_lr, probitMod_work_lr), 
               title = "Rural sample probit regression results (lagged)", 
               label = "fullResults6",
               header = F,
               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

```

##Robustness Checks 

The following results show a different income specification, the inclusion of the extreme poverty variable instead of the log of income.

```{r probit regression total sample extr, echo = F, message = F}

probitMod_assistx <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))  

probitMod_registerx <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

probitMod_workx <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

```

```{r generate probit regression table extr, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assistx, probitMod_registerx, probitMod_workx), 
               title = "Full sample probit regression results (poverty measure)", 
               label = "fullResults8",
               header = F,
               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))

```

The following results show the linear regression models.

```{r linear regressions total sample, echo = F, message = F}

model_assist_t <- lm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)

model_register_t <- lm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)

model_reason_t <- (lm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))

```

```{r generate regression table, echo = F, results = "asis"}
stargazer::stargazer(list(model_assist_t, model_register_t, model_reason_t), 
               title = "Full sample linear regression results", 
               label = "fullResults11",
               header = F,
               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
```

#Discussion
